\chapter{Arms and Armour}
\label{chap:arms}

\begin{table}[ht!]
	\centering
	\caption{Close-combat Weapons. Note that L signifies \textlf{lethality}, A is \textlf{aim}, D is \textlf{deflect}, and P is \textlf{power}. Pen indicates the \textlf{penetration} effect and MD is \textlf{massive damage}.}
	\label{tab:weps}
	\begin{tabular}{|l|l|l|l|l|l|l|l|}
		\hline
		Name & Cost & Hands & L & A & D & P & Special\\ [0.5ex]
		\hline 
		\textbf{Daggers} & & & & & & & \\
		\hline
		Dagger & 5 s & 1 & N & - & - & - & Small, Throw 1\\
		Stiletto & 10 s & 1 & N & - & - & - & Small, Rending \\
		\hline 
		\textbf{Swords} & & & & & & & \\
		\hline
		Arming Sword & 20 s & 1 & N & - & +1 & - & \\
		Long Sword & 60 s & 2/1 & N & - & +1 & +1/- & \\
		Falchion & 40 s & 1 & N & - & - & +1 & - \\ 
		Rapier & 50 s & 1 & N & - & +1 & - & Rending \\ 
		Sword-breaker & 10 s & 1 & N & - & +1 & - & Disarm\\
		Greatsword & 75 s & 2 & N & - & +1 & - & MD\\
		Zweihander & 2 g & 2 & C & - & - & +1 & Cumbersome\\ 
		\hline 
		\textbf{Axes} & & & & & & & \\
		\hline
		Battle Axe & 25 s & 1 & N & +1 & - & - & - \\
		Bearded Axe & 40 s & 1 & N & +1 & - & - & Disarm \\
		Throwing Axe & 10 s & 1 & N & +1 & - & - & Throw 2 \\
		Great Axe & 70 s & 2 & N & +1 & - & - & MD \\
		Long Axe & 70 s & 2 & N & +1 & - & - & MD, Reach \\
		Bearded Long Axe & 90 s & 2 & N & +1 & - & - & MD, Disarm, Reach \\
		\hline
		\textbf{Pole Weapons} & & & & & & & Trip \\
		\hline
		Javelin & 5 s & 1 & N & - & - & - & Rending, Throw 2\\
		Quarter Staff & 10 c & 2 &  N & - & - & - & Disarm,Reach\\
		Spear & 10 s & 2/1 & N & - & - & -/- & Rending, Reach/- \\
%		Pike & 25 s & 2 & N & - & - & +1 & Rending, Reach\\
		Glaive & 40 s & 2 & N & - & - & +1 & Reach \\
		Halberd & 1 g 50 s & 2 & N & - & - & - & Multi-edged, Reach\\
		Lucerne Hammer & 1 g & 2 & N & - & - & - & Multi-edged, Reach\\ 
		Ranseur & 40 s & 2 & N & - & - & - & Rending, Disarm \\ 
		Partisan & 40 s & 2 & N & - & +1 & - & Rending\\
        Pole-axe & 1 g & 2 & N & - & - & - & Multi-edged \\
		Lance & 1 g & 2 & N & - & - & - & Rending, Reach\\
		\hline
		\textbf{Blunt Weapons} & & & & & & & \\
		\hline
		Cudgel & 20 c & 1  & N & - & - & - &  \\
		Club & 2 s & 2 & N & - & - & +1 & Pen 1 \\
		War Hammer & 40 s & 1 & N & - & - & - & Pen 2 \\ 
		Great Hammer & 60 s & 2 & N & - & - & - & Pen 2, MD \\
		Maul & 70 s & 2 & C & - & - & - & Cumbersome, Pen 1 \\
		Mace & 5 s & 1 & N & - & - & - & Pen 1 \\
		\hline
		\textbf{Extended Weapons} & & & & & & & \\
		\hline
		Flail & 50 s & 1 & N & - & - & +1 & Flail Fail \\ 
		Grand Flail & 1 g & 2 & N & - & - & +2 & Flail Fail, Trip, Reach \\
		Whip & 5 s & 1 & N & - & - & -1 & Trip, Disarm, Range 1\\
		Chain & 25 s & 2 & N & - & - & - & Trip, Disarm, Reach\\  
%		Garotte Wire & 50 s & 2 & C & - & Sneak Attack only\\
		\hline
	\end{tabular}
\end{table}
\begin{table}[ht!]
	\centering
	\caption{Ranged Weapons. Note that L signifies \textlf{lethality} and P is \textlf{power}. Pen indicates the \textlf{penetration} effect and MD is \textlf{massive damage}.}
	\begin{tabular}{|l|l|l|l|l|l|l|l|}
		\hline
		Name & Cost & Hands & Range & Reload & L & P & Special\\ [0.5ex]
		\hline
		\textbf{Bows} & & & & & & & \\
		\hline
		%Hunting Bow & 5 s & 2 & 2 & 1 & N & +1 & - \\
		Short Bow & 20 s & 2 & 1 & - & N & - & -\\
		Long Bow & 60 s & 2 & 3 & 1 & N & - & MD \\
		Recurve Bow & 1 g & 2 & 2 & - & N & + 1 & -\\
		\hline
		\textbf{Cross Bows} & & & &  & &  & \\
		\hline
		Hand Crossbow & 50 s & 1 & 1 & - & N & - & Small \\ 
		Light Crossbow &  50 s & 2 & 2 & 2 & N & +1 & MD\\
		Repeater Crossbow & 70 s & 2 & 1 & 1 & N & - & - \\  
		Heavy Crossbow & 3 g & 2 & 3 & 3 & C & +1 & MD \\
		\hline
		\textbf{Black Powder} & & & & & & & Powder-Shot \\
		\hline
		Pistol & 1 g & 1 & 1 & 1 & N & - & Pen 1, MD \\
		Musket & 3 g & 2 & 3 & 2 & N & - & Pen 2, MD \\
		Blunderbuss & 2 g & 2 & 1 & 2 & N & - & Burst 1, Cone, MD\\
		Bomb & 50 s & 1 & 1 & - & N & - & Burst 1, Throw, Radius 0, MD \\
		\hline 
	\end{tabular}
	\label{tab:range-weps}    
\end{table}

\begin{table}[ht]
	\centering
	\caption{Basic Armour}
	\begin{tabular}{|l|l|l|l|}
		\hline
		Name  & Cost & Type & Toughness\\
		\hline
		Light Armour & & &  \\
		\hline
		Gambeson & 20 s & L & +1  \\
		\hline
		Medium Armour & & &  \\
		\hline
		Mail hauberk and gambeson & 2 g & M & +3   \\ 
		Brigandine and mail  & 4 g & M & +4  \\
		\hline
		Heavy Armour & & &  \\
		\hline
		Brigandine and plate & 6 g & H & +5 \\
		Full Plate  & 10 g & H & +6  \\
		\hline
		Barding (Horse)& 6 g & H & +4 \\
		\hline
	\end{tabular}
	\caption{Shields}
	\begin{tabular}{|l|l|l|l|}
		\hline
		Name & Cost & Deflect & Special\\   
		\hline
		Shield & 30 s & +2 & - \\
%		Round Shield & 50 s & +2 & - \\
%		Tower Shield & 2 g & +3 & Adamant, Cumbersome \\
		\hline
	\end{tabular}
\end{table}





\section{Weapon Special Rules}

\subsection{Multi-Edged}
A weapon like this has many different naughty ends. Which bit is being used must be decided before rolling each attack. A halberd can \textlf{trip} with its hook, slash with the axe-blade (+ 1 \textlf{Power}) or stab with the spear point (\textlf{Rending}). The Lucerne Hammer is similar but it has \textlf{Penetration} 2 when the hammer head is used. A pole-axe has a heavy axe blade with \textlf{massive damage}, a hammer-head with \textlf{penetration} 2, and a spear tip with \textlf{rending}.

\subsection{Flail}
A flail is a one-handed weapon consisting of a short handle attached to a heavy mace head on the end of a chain. The extension of the chain grants the weapon great power for a single-handed weapon. However, the chain can get tangled or caught on an opponent. Thus, any \textlf{Critical success} \textlf{Deflects} by targets result in the wielder being forced to untangle the flail before it can be used to attack, which results in a \textlf{Moment of Weakness}. 

\subsection{Lance}
A lance gains bonus \textlf{lethality} during a mounted charge action.

\subsection{Bows}
Bows are shooting-type weapons. Firing two arrows together incurs an \textlf{edge} penalty to \textlf{Aim} but adds an extra projectile to the shot.

%\subsubsection{Long Bows}
%These huge bows are difficult to draw, requiring a mighty arm to get the best results. These cannot be used without a positive-definite \textlf{Might} score ($>0$).

\subsection{Crossbows}
All crossbows are shooting weapons. Unless specially modified, these cannot fire two bolts at once.

\subsubsection{Hand Crossbow}
These diminutive weapons are designed to be easily concealed and to unloaded at close range into an unsuspecting victim. As such, they are \textlf{small} and fast firing but are still tricky to reload when used in pairs (requiring a 1 action \textlf{Reload}).

\subsection{Powder-Shot}
Black powder weapons are powerful but quite dangerous to their wielder. As such, scoring a raw $5-$ on \textlf{aim} means the weapons firing goes dangerously awry. This causes the user a single damage roll (use \textlf{power} +2 and \textlf{lethality} N). In the case of the bomb, treat it as though it hit everyone inside the same area as the character (including themselves) and resolve hits as normal.

\subsubsection{Blunderbuss}
This weapon fires a scatter of shot in a cone, hitting up to 2 targets in a single combat area. The user can also declare using it point blank within range 0, then it hits one target only but has \textlf{Burst} +1.





\section{Armour Special Rules}
\label{sec:armspec}

\subsection{Armour Weights}
Armour is listed with a weight appropriate to medium sized armour, that is, armour made for medium sized creatures to wear. Armour for smaller creatures weighs half as much and armour for Large creatures weighs twice as much, larger armour weigh six times as much as medium armour.


\subsection{Putting On and Removing Armour}
Even heroes seldom sleep in full-plate armour, so there are then times when the speed at which a character puts on or removes armour might matter. It requires four action points to put on \textlf{light armour} or two to strap on a shield. However, it takes 1 minute to put on \textlf{medium armour} but it takes 5 minutes to put on \textlf{heavy armour} and requires that the wearer has assistance in putting it on. \textlf{Medium} or \textlf{light armour} can be put on in a rush, doing so means that armour might not be precisely adjusted in order function to it's full effect, this takes only four action points for medium armour or two action point for \textlf{light armour}, but reduces the \textlf{Toughness} of the armour by 1 while it is worn in this sloppy fashion.

Removing \textlf{light armour} takes 1 minute, \textlf{medium armour} takes 2 minutes, and \textlf{heavy armour} takes 5 minutes to remove.


\subsection{Shields}
\label{sec:shields}
Shields add a + 2 bonus to the bearer's \textlf{Deflect} score. 


\section{Weapon Enhancements}
A weapon may have one quality enhancement and one other enhancements. However, weapons cannot have more than one that is labelled `Exclusive'.

%\subsection{Crank and Magazine}
%(Crossbow only - Exclusive) The weapon is fitted with a complex mechanical crank and ammunition store, allowing for rapid firing. This allows crossbows with a \textlf{Reload} rule to fire twice before having to reload. Other crossbows may gain \textlf{Burst} + 1 when firing at the cost of a \textlf{Reload} 1 action next turn. Cost: 8 g.

\subsection{Fine}
(Quality enhancement) This grants the item + 1 \textlf{power}. This doubles the cost of the item. 

\subsection{Masterly}
(Quality enhancement) This grants the item + 1 \textlf{Power} and \textlf{aim}. This quadruples the cost of the item.

\subsection{Expanded Bolt Rack}
(Crossbow only) The weapon is fitted with expanded armatures and space for two bolts. It can thus fire two bolts with a single trigger release. This grants \textlf{Burst} + 1 on attacks made with the crossbow. Cost: Base cost x 3 + 1 g. 

\subsection{Special Payload}
(Crossbow only) The armatures and bolt track of the bow are modified to carry larger and heavier bolts. This configures the crossbow to fire specialised ammunition that can carry explosives (bomb from infusary - Section~\ref{sec:alch}), chains/ropes/nets, or poisons and other chemicals that spray in a radius of 0 around the impact point. Cost: 5 g.

\subsection{Experimental Breach}
(Black Powder Weapons only) The weapon can be loaded through a opening in the barrel, rather than down the muzzle. This reduces the \textlf{Reload} time of the weapon by one rank (reduce paired pistols to \textlf{Reload} 1). However, if the user scores a \textlf{Critical Failure} with the weapon roll 3d6 in addition to normal \textlf{Black Powder} effects. On a score of $8-$ the weapon needs 50 s worth of repairs before it works again, on $9+$ it jams requiring 2 action points be spent to clear the jam. Cost: 10 g.

\subsection{Extra Barrel}
(Black Powder Weapons only) The weapon is fitted with an extra barrel. The barrels can be fired together adding an \textlf{edge} bonus to damage rolls, or separately, in which case the gun may be fired twice before re-loading. Cost: 3 g.




\section{Armour Enhancements}
Armour may have one quality enhancement and only one other enhancement.

\subsection{Fine}
(Quality enhancement) This grants the item + 1 Toughness. This doubles the cost of the item.

\subsection{Masterly}
(Quality enhancement) This grants the item + 2 Toughness. This quadruples the cost of the item.

\subsection{Thick Padding}
A hefty layer of cloth padding soaks up the impact of musket fire. This reduces crushing lethality \textlf{Black Powder} weapons to normal \textlf{lethality} but adds an \textlf{edge} penalty on \textlf{Athletics} and \textlf{Stealth}. Cost: 50 s.

\subsection{Intricate Scale-work}
Fine layers of scaled plates distribute impact. This incurs a -1 \textlf{Burst} penalty (minimum 1 damage roll) to any attacks made against the wearer. Cost: 12 g. \textlf{Medium} or \textlf{heavy armour} only.

\subsection{Heavily Reinforced}
Nothing like just adding lots of extra steel. Adds the \textlf{Adamant} rule. Cost: 10 g. \textlf{Medium} or \textlf{heavy armour} only.

\subsection{Cunning Strappage}
Careful use of straps and weight distribution makes the armour feel so light you could dance in it (this is still not recommended). This negates a chosen \textlf{edge} penalty incurred by the armour. Cost: 6 g. \textlf{Medium} or \textlf{heavy armour} only.

\subsection{Knuckle-Blades}
Punching people is easier with steel blades on the knuckles of your gloves. This granst + 1 \textlf{power} to unarmed damage rolls. Cost: 50 s.

\subsection{Armour Spikes}
Sharpened protrusions or heavy studs set in the surface of the armour can make attacking the wearer a real hazard. This can be applied to any \textlf{Medium} or \textlf{Heavy armour} piece, enemies grappling with the wearer have an \textlf{edge} penalty to \textlf{Grapple} checks. Cost 20 s.



\chapter{Other Equipment}

\section{Currency}
There are 10 copper pieces to a silver, and 100 silver to a gold.

\section{Cost}
This cost reflects an average price that the item would be purchasable for, from a merchant. This cost can go up and down dependent on how well a player haggles and on the circumstances of the city and/or merchant; that is, if something is in shortage it costs a lot more, or if the merchant is desperate to sell, the price goes down. A player can sell an item for a price dependent on the condition of the item, a merchant will offer him half the normal cost for a used item in good condition.

\section{Equipment}
\subsection{Tools}

\subsubsection*{Bandages}
Required by a healer to use his healing skill, each use of the skill consumes a bandage.

\subsubsection*{Infuser's tools}
A complex piece of equipment that allows an infuser to purify and distil various liquids, this is required for the creation of infusions (see Section~\ref{sec:alch}).

\subsubsection*{Glass Vial}
A simple glass vial with a stopper, used to hold liquids, this is needed to hold any brew created by an infuser (see Section~\ref{sec:alch}).

\subsubsection*{Lock-pick}
A small piece of cunningly bent metal that can be used to open a lock with a Mechanical skill check opposed to the difficulty of the lock, on a failure the lock-pick breaks.

\subsubsection*{Musical Instrument}
A simple instrument for making music.

\subsubsection*{Wood Axe}
An axe for chopping up wood.

\subsubsection*{Master-Crafted Tools}
These are finely crafted versions of any type of tool, this adds an \textlf{edge} bonus to any rolls made using the tools.

\subsection{Adventuring Gear}

\subsubsection*{Backpack}
A large but otherwise conventional satchel.

\subsubsection*{Barrel}
A large wooden barrel, bound with iron rings. The barrel can hold about 50 litres.

\subsubsection*{Basket}
A simple wicker basket for carrying small loads.

\subsubsection*{Candle}
A small candle that that produces \textlf{low} light in a 6 m radius, this candle can burn for up to 3 hours.

\subsubsection*{Canvas}
A sheet of water-proof canvas. The price depends on the size of the sheet.

\subsubsection*{Chain}
A metal chain made up of heavy steel links, the chain is strong enough to support very large weights, up to 500 kg.

\subsubsection*{Crowbar}
A simple steel crowbar used for levering open doors or hinges. Using a crowbar adds an \textlf{edge} bonus to rolls made using your Might to force open any hinged container or door.

\subsubsection*{Firewood}
Wood for keeping a fire burning for one day.

\subsubsection*{Fishing Tackle}
A fishing line, a hook, sinkers and lures. This can be used to attempt to catch fish with the \textlf{Survival} skill, the difficulty of this is dependent on the speed of the water and type of the fish.

\subsubsection*{Grappling Hook}
A heavy iron hook used for scaling near vertical surfaces. This hook will hold the attached rope in place after being successfully thrown to the point you wish to climb to. The hook also makes an effective weapon.

\subsubsection*{Hammer}
A simple hammer for hitting stuff and pitching tents.

\subsubsection*{Ladder}
A length of ladder which can be climbed without needing an Athletics check.

\subsubsection*{Lantern}
A hooded lantern used for projecting \textlf{full} light directionally, up to 13 m from the bearer. This uses 0.5 kg of oil to burn for 24 hours.

\subsubsection*{Lantern oil}
Enough oil to power a standard lantern for 24 hours. Highly flammable.

\subsubsection*{Mirror}
A simple shiny surface (glass or metal) that reflects light. Useful for looking around corners.

\subsubsection*{Needle}
A sharp needle for use in sewing or stitching wounds.

\subsubsection*{Paper}
A single sheet of paper for writing on.

\subsubsection*{Pick}
A simple pick used for breaking earth or skulls.

\subsubsection*{Pole}
A length of wooden or metal pole.

\subsubsection*{Pot}
A cooking pot, about 30 cm in diameter.

\subsubsection*{Quill}
A feather quill used for writing.

\subsubsection*{Rope}
A length of rope capable of supporting the weight of three people ($\sim$250 kg).

\subsubsection*{Sack}
A large rough cloth bag.

\subsubsection*{Spade}
A digging implement (can be used to hit people as well).

\subsubsection*{Spyglass}
A telescope capable of magnifying things from up to a mile away.

\subsubsection*{Tent}
A canvas tent to keep you dry at night, this can house up to two man-sized creatures.

\subsubsection*{Tinder Box}
A small box containing fast lighting twigs, used to rapidly start a fire.

\subsubsection*{Torch}
Provides \textlf{full} light up to a 6 m radius and \textlf{low} light within 12 m. The torch can burn for 4 hours.

\subsubsection*{Trail Rations}
Generally dried meat, bread and cheese. Long lasting food-stuffs to nourish a traveller.

\subsubsection*{Water-Skin}
A hide bag used for carrying water, this carries water for one person for 5 days.

\subsection{Mounts}
\subsubsection*{Pony}
A small hard working pony, capable of carrying an full grown small creature or the child of a man sized creature. A pony has a combat movement distance of 2. Such a creature has Might -1 but can carry weights of up to 100 kg while still being to walk or trot all day with no appreciable strain. Carrying weights of 150 kg or more risks injuring the horse if done for prolonged periods. This horse is not trained for war and may panic, it has Deflect -2 and Toughness 8.

\subsubsection*{Light Horse}
A swift if none too tough horse, a Light Horse has a combat movement distance of 4. A light horse can gallop up to 2 hours a day at a speed of five times faster than a man, any longer risks injuring the horse. This horse can only move a distance of 2 per movement action while carrying a heavily-armoured rider. A light horse has Might 0 but can carry weights of up to 80 kg while still being to walk or trot all day with no appreciable strain. Carrying weights of 130 kg or more risks injuring the horse if done for prolonged periods. This horse is not trained for war and may panic, it has Deflect 0 and Toughness 9.

\subsubsection*{Cart Horse}
A heavy farm horse capable of carrying heavy loads and working all day, however, it is not capable of maintaining very rapid speeds. A Cart Horse has a combat movement distance of 2. A cart horse has Might +2 but can carry weights of up to 200 kg while still being to walk or trot all day with no appreciable strain. Carrying weights of 300 kg or more risks injuring the horse if done for prolonged periods. This horse is not trained for war and may panic, it has Deflect -2 and Toughness 11.

\subsubsection*{Heavy Horse}
A strong and reasonably quick horse. This horse can carry a man in heavy armour while moving at full speed. This horse has a combat movement distance of 3. Such a creature has Might +1 but can carry weights of up to 120 kg while still being to walk or trot all day with no appreciable strain. Carrying weights of 200 kg or more risks injuring the horse if done for prolonged periods. This horse is not trained for war and may panic, it has Deflect -2 and Toughness 10.

\subsubsection*{Destrier (War Horse)}
A heavy horse trained for battle, the destrier is by far the largest breed of horse, standing over 2 metres at the shoulder. Its war training makes it a savage weapon in its own right and it has close-combat Aim +1, Deflect 0, Toughness 10, Power +3, and combat movement range 3. A destrier's damage rolls all have Massive Damage. A destrier can carry weights of up to 200 kg while still being to walk or trot all day with no appreciable strain. Carrying weights of 300 kg or more risks injuring the horse if done for prolonged periods.

%\subsubsection*{Dire Wolf}
%A ravening grey wolf ridden by goblins, taming this creature usually only requires the death of a couple of goblins. A wolf can carry one small creature and has a combat movement distance of 2. A dire wolf has Aim and Deflect +1, Toughness 7, and its attacks make two damage rolls each with Power +2. Such a mount is not heavy enough to Trample over enemies effectively. A wolf has Might +2 but can carry weights of up to 60 kg while still being to walk or trot all day with no appreciable strain. Carrying weights of 100 kg or more risks injuring the horse if done for prolonged periods.
%
%\subsubsection*{War Boar}
%A foul-smelling, ill-tempered beast ridden commonly by the orcs. These creature's ferocious charge causes a devastating impact, inflicting an additional hit at Power +5 when making an attack after moving. This creature has Aim +1, Deflect -3, Toughness 12, and its attacks make two damage rolls each with Power +3. The boar has Might +3 and the rider must make an opposed Might check if he wishes to order the boar to do anything. Otherwise the boar will move towards the nearest edible or offensive object, and eat or kill it as appropriate. The boar's combat movement distance is 2.

\subsubsection*{Simple Saddle}
A rough worked leather saddle, needed for sitting astride your mount. A saddle is fastened with girth buckles and has attached stirrups.

\subsubsection*{War Saddle}
A heavy saddle needed to support a fully-armoured knight, this saddle is designed also to allow the knight to wield a lance one handed.

\subsubsection*{Racing Saddle}
A saddle built for race horses, it is light weight and fine craftsmanship add 1 to the horses' combat movement distance.

\subsubsection*{Tack}
Reins, bit, bridle and other equipment necessary for riding a mount. Riding bareback increases the difficulty of all ride checks by 2.

\subsubsection*{Feed}
Food for a mount for one day.

\chapter{Equipment Tables}
\begin{table}[ht]
	\centering
	\caption{Tools}
	\begin{tabular}{|l|l|}
		\hline
		Item & Cost\\ [0.5ex]
		\hline
		Bandages (5)& 2 s\\
		Infuser's tools & 1 g\\
		Glass Vials (5)& 5s\\
		Lock-picks (5) & 10 s\\
		Musical Instrument & 50 s\\
		Wood Axe & 5 s\\		
		\hline
		Master-Crafted Tools & +3 g\\
		\hline
	\end{tabular}
\end{table}

\begin{table}[ht]
	\centering
	\caption{Adventuring Gear}
	\begin{tabular}{|l|l|l|}
		\hline
		Item & Cost & Weight\\ [0.5ex]
		\hline
		Backpack & 5s & 1 kg\\
		Barrel & 10 s & \\
		Basket & 2 s & 0.2 kg\\
		Candle & 10 c & -\\
		Canvas (per sq. m) & 1 s & 4 kg\\
		Chain, 3 m & 10 s & 10 kg\\
		Crowbar & 25 s & 5 kg\\
		Firewood, 1 day & 20 c & 5 kg\\
		Fishing Tackle & 1 s & -\\
		Grappling Hook & 10 s & 5 kg\\
		Hammer & 5 s & 4 kg\\
		Ladder, 3 m & 10 s & 10 kg\\
		Lantern & 15 s & 1 kg\\
		Lantern oil & 5 s & 0.5 kg \\
		Mirror & 5 s & -\\
		Needle & 1 s & -\\
		Paper 1 sheet & 1c & -\\
		Pick & 5 s & 8 kg\\
		Pole, 3 m & 1 s & 5 kg\\
		Pot & 10 s & 0.5 kg\\
		Quill Pen & 10 c & -\\
		Rope, 3 m & 1 s & 0.2 kg\\
		Sack & 10 c & 0.2 kg\\
		Spade & 5 s & 4 kg\\
		Spyglass & 1 g & 1 kg\\
		Tent & 20 s & 3 kg\\
		Tinder Box & 5 s & -\\
		Torch & 2 c & 0.1 kg\\
		Trail Rations 1 day & 10 c & 0.1 kg\\
		Water-skin (full) & 1 s & 3 kg\\		
		\hline
	\end{tabular}
\end{table}

\begin{table}[ht]
	\centering
	\caption{Mounts}
	\begin{tabular}{|l|l|}
		\hline
		Item & Cost\\ [0.5ex]
		\hline
		Pony & 75 s\\
		Light Horse & 1 g 20 s\\
		Cart Horse & 4 g\\
		Heavy Horse & 4 g\\
		War Horse & 10 g\\
%		Wolf & 2 g\\
%		Boar & 3 g\\
		\hline
		Simple Saddle & 10 s\\
		War Saddle & 50 s\\
		Racing Saddle & 50 s\\
		Tack & 20 s\\
		Feed, 1 day & 5 c\\
		\hline
	\end{tabular}
\end{table}

\begin{table}[ht]
	\centering
	\caption{Clothes}
	\begin{tabular}{|l|l|}
		\hline
		Item & Cost\\ [0.5ex]
		\hline
		Peasant & 5 c\\
		Scholar & 10 s\\
		Cleric & 5 s\\
		Cold Weather & 20 s\\
		Noble & 2 g\\
		Merchant & 50 s\\
		\hline
	\end{tabular}
\end{table}
