\section{Aether manifestation}
\label{sec:spells}

To represent a character's development in aethermancy they can learn new manifestations by paying the experience point cost listed in square brackets after the name. No manifestations can be learned unless the character has \textlf{Skill proficiency: aethermancy}. All manifestations cost 1 action point unless otherwise specified. \textlf{persistent} manifestations are marked with (P) after their experience cost.

%\subsection{Accusation [2]}
%You accuse a target creature of some misdeed, all nearby creatures become suspicious that the accusation is true if they fail a \textlf{Resist(R)} check.
%\subsubsection{Augment: Denouncement [2]}
%The target of Accusation is subject to the \textlf{Cripple} effect for 1 round and 1 additional round per level of \textlf{critical failure} on \textlf{resist}.
%
\subsection{Aether siphon [2]}
An aethermancer can establish a resonance with another creature such that the aether is stripped from their body. For 2 action points (but no aether points) a chosen target within range 0 (5 m) must \textlf{resist(C)} or lose an aether point. If the target has no aether points, it loses an \textlf{endurance} instead. If the aethermancer wins the \textlf{resist} check, they gain 1 aether point.


\subsection{Aetheric blast [2]}
\label{spell:sorc-blast}
This costs 1 action point and projects your aether out in a searing jet. A chosen target within range 1 (15 m) must \textlf{deflect}, if they fail then they must \textlf{resist(R)} or suffer damage with normal \textlf{lethality} (\textlf{critical success} for the aethermancer on \textlf{resist} increments damage severity). You have \textlf{Edge} on the \textlf{resist} roll. 
\subsubsection{Augment: Resounding blast [2]}
Aetheric blast now inflicts \textlf{knocked down} if it causes damage.


\subsection{Aetheric wall [3] (P)}
For 1 action point you manifest a wall of shimmering aether around a radius of 5 m (1 combat area) within a range of 2 (25 m). Passing through the wall costs any creature 1 action point and requires a \textlf{resist(M)} check, failure means they cannot traverse the wall and suffer damage with normal \textlf{lethality} (\textlf{critical success} for the aethermancer on the \textlf{resist} increments the damage severity). This lasts for 10 minutes and is a \textlf{persistent} manifestation.


\subsection{Animal aether [2]}
Every living creature hums with an aetheric field. By carefully tuning your own aetheric field into resonance you can hear the mood and emotions of animals.
\subsubsection{Augment: Animal resonance [2]}
Your mastery of the aetheric fields of animals allows you to communicate emotions to them as well as attempt to \textlf{persuade} them.


\subsection{Animate homunculus [2]}
You can shape a small creature out of any given material (30 cm is the maximum size), this creature is then imbued with aether. This grants it a kind of limited life while directly controlled by you (this requires your full concentration). The creature will only act when controlled, otherwise it remains inert. A homunculus has no skill proficiencies and cannot inflict damage in combat.
\subsubsection{Augment: Automaton [4]}
Your homunculi are always animated and will obey simple verbal commands. You may have 1 + \textlf{Wit} homunculi active at once. 
\subsubsection{Augment: Golem [5]}
The maximum size, for a single homunculus, is increased to medium (man-sized) creatures, these can wield weapons and make unarmed attacks in combat. 


\subsection{Arcing aether [3]}
You infuse your aether into a resonance with elemental lightning. For 2 action points, aetheric lighting leaps from your body and strikes a target within range 2 (25 m). The target must \textlf{resist(R)} or suffer damage with normal \textlf{lethality}. The lightning then leaps to a second target within range 1 of the first. It will only leap to the aethermancer's allies if no enemies are in range. 
\subsubsection{Augment: Resonant arcing [2]}
The lightning gains an extra leap every time it causes damage.
\subsubsection{Augment: Paralytic arcs [2]}
Victims damaged by this manifestation are \textlf{immobilised} for 1 round.

\subsection{Befuddle [2]}
A discordant aether infusion removes your target's ability to distinguish friends from foes. If the target fails a \textlf{resist(W)} check, they regard all creatures as hostile and dangerous. This means they can panic, run away, or lash-out at anything nearby. Choose combat targets and behaviour for the victim randomly. This lasts up to 1 minute. Each round after the first, the victim may re-attempt to \textlf{resist}.


\subsection{Bend light [3] (P)}
\label{spell:bend-light}
Aether flows from your fingers creating an aetheric lattice that bends light. At a distance up to range 1 (15 m), you can create a stationary illusion up to medium size. Anyone looking at this effect must \textlf{resist(W)} to decide if they are fooled. This is a \textlf{persistent} effect. The illusion is only visual, it makes no sounds, smells, and cannot be touched. 
\subsubsection{Augment: Major illusion [2]}
You can bend light with such dexterity that you can create illusions up to 5 m in size. 


\subsection{Blinding ray [2]}
\label{spell:blind-ray}
You pour aether into a light source you touch, focusing its illumination into a searingly bright beam. This \textlf{blinds} a victim that fails a \textlf{resist(R)} check, additionally the normal lighting from the light source is removed for 1 round but the area adjacent to the target is fully illuminated for this time. The blindness lasts 1 additional round per level of \textlf{critical success} for the aethermancer on \textlf{resist}.


\subsection{Clinging aether [3]}
For 1 action point you blast a 5 m radius (1 combat area) within a range of 1 (15 m) with crackling aether that anchors everything to the earth. All creatures in the area must \textlf{resist(C)} or be \textlf{immobilised} for 1 round (1 additional round per level of aethermancer \textlf{critical success} on the \textlf{resist} check).


\subsection{Compel [3]}
You infuse aether into a target by touch, forcing them to make an action chosen by you if they fail a \textlf{Resist(W)} check. This fails if the action would harm the target. This is a \textlf{persistent} effect. Specify one extra action per level of \textlf{critical success} for the aethermancer on \textlf{resist}. The victim is aware that they being forced to act.
\subsubsection{Augment: Dominate [3]}
Your control is so strong that Compel will not fail if the action would be harmful to the victim itself.
\subsubsection{Augment: Project aether [2]}
You can now manifest this power at range 1.


\subsection{Damping field [2] (P)}
\label{spell:arcane-armour}
For 1 action point you envelope yourself in a high-pressure field of aether that depletes the force of incoming attacks. This manifests as a shimmering in the air around you and increases your \textlf{toughness} by your \textlf{wit} score. You cannot manifest this while wearing armour, as this disrupts the widespread projection of your aether. This manifestation lasts for 10 minutes and is a \textlf{persistent} effect.
\subsubsection{Augment: Hardened aether [3]}
Damage rolls against you suffer an \textlf{edge} penalty while this manifestation endures.


\subsection{Earthquake [4] (P)}
You flood the ground with aether, making it roll and buck as though it was alive. For 2 action points you make the ground violently shake. All creatures adjacent to you must \textlf{resist(M)} or be knocked down. This can be maintained as a \textlf{persistent} effect for up to 1 minute. 
\subsubsection{Augment: Project aether [2]}
You can now manifest this power at range 2 (25 m), instead of shaking the ground adjacent to you.


\subsection{Eyes of aether [2]}
You can alter your senses to perceive flows of the aether. This renders you blind, instead you view the world as the flow of the aether. This costs no aether points to use and may be cancelled at any time.


\subsection{Fiery surge [3]}
You can fuel a flame via aether infusion, greatly increase the intensity of an existing fire within range 1 (15 m). Any creature that makes contact with such a flame must \textlf{resist(C)} or suffer damage with crushing \textlf{lethality}. This fire burns for 1 minute.


\subsection{Fear [2]}
You infuse a target, within range 1 (15 m), with aether, causing a cloud of fear to pass across the target's mind if they fail a \textlf{resist(R)} check. Such a target cannot approach the aethermancer or remain adjacent to them for 1 minute. Each round after the first the victim may re-attempt the \textlf{resist(R)}. 
\subsubsection{Augment: Terror [3]}
You have learned to afflict the mind with absolute terror. The target of fear suffers an \textlf{edge} penalty to all rolls while this effect persists. 



\subsection{Flare [2]}
\label{spell:flare}
Projecting aether suddenly into a flame within range 1 (15 m), you cause it to burst into in blinding white flash. Anyone who can see the flare is \textlf{blind} for 1 round and 1 additional round per level of \textlf{critical failure} on \textlf{resist(M)}. The explosion of the flare itself is harmless.
\subsubsection{Augment: Burning flare [3]}
Flare ignites all adjacent creatures, who suffer damage with normal \textlf{lethality} if they failed \textlf{resist(M)} (\textlf{critical success} for the aethermancer on \textlf{resist} increments damage severity).


\subsection{Fireball [4]}
This costs 2 action points. You condense your aether into an explosive fireball that you can hurl at a 5 m radius (1 combat area) within range 2 (25 m). Creatures within the blast must \textlf{deflect} or make a \textlf{resist(R)} check. Failure on the latter inflicts damage with normal \textlf{lethality} (\textlf{lethality} scales with levels of \textlf{critical success} for the aethermancer). This manifestation sets all flammable material in the radius on fire.
\subsubsection{Augment: Blaster Master [3]}
The fiery explosion burns with voracious intensity. This manifestation gains the \textlf{Heavy weapon} 1 effect.


\subsection{Freezing storm [4] (P)}
For 2 action points you choose a region of radius 5 m radius (1 combat area) within a range of 2 (25 m) and rain a shower of freezing ice shards upon it. Creatures within the area suffer a hit with normal \textlf{lethality}. For 10 minutes the area is \textlf{rough terrain}, anyone entering the area suffers a hit with normal \textlf{lethality}. This is a \textlf{persistent} effect.
\subsubsection{Augment: Sub-zero [2]}
The cost of this manifestation is reduced by 1 action point.


\subsection{Glacial ray [3]}
This costs 2 action points. A ray of freezing aether strikes a target within range 1 (15 m). They must \textlf{deflect} and then \textlf{resist(R)} if they fail. Failure on the second check means they are \textlf{immobilised} and \textlf{vulnerable} until damaged. 
\subsubsection{Augment: Creeping frost [2]}
Glacial ray applies its effects to all creatures adjacent to the primary target.
\subsubsection{Augment: Deep-freeze [2]}
After the effect of Glacial ray ends the ice shatters causing all creatures adjacent to the target take a hit with normal \textlf{lethality}.


\subsection{Guided motion [3]}
Aether streams out from you to enhance the movements of another body within range 1 (15 m). This confers an \textlf{edge} bonus to an allied target's next \textlf{Athletics} or \textlf{deflect} check.
\subsubsection{Alternate: Inhibit motion [2]}
Guided motion can target an enemy and confer an \textlf{edge} penalty instead if they fail a \textlf{resist(M)} check.


\subsection{Ignite [2]}
\label{spell:ignite}
You infuse your aether into generating a resonance between your target and the element of fire. This manifestation can be used on a chosen target within range 1 (15 m). The fire created by Ignite also suffers from normal physical restrictions., i.e. You may not set fire to a creature unless it is naturally flammable, covered in oil, or circumstantially vulnerable. Materials like metals can be heated by this effect to burn their bearer. Setting fire to the clothes of a foe does no great harm to him (it might cause weak enemies to panic).


\subsection{Illuminate [2]}
\label{spell:illuminate}
You infuse an object with aether, causing it to resonate with the aetheric field of light. A single small object you touch begins to emit a soft aetheric glow. This provides low-light illumination over a single combat area (around a 5 m radius). This lasts for up to 6 hours.


\subsection{Iron arm [4] (P)}
You infuse a creature with aether that resonates with the musculature of their body. A chosen creature within range 1 (15 m) gains +1 \textlf{power}. This is a \textlf{persistent} effect that lasts up to 1 minute.


\subsection{Mimic sound [2]}
You subtly leak aether into the air, causing it to vibrate in a chosen pattern. This allows you to produce any sound you can imagine, at volumes between a whisper and a shout. The convincingness of this is decided by results of listeners' \textlf{resist(W)} checks.


\subsection{Orb of aether [3] (P)}
\label{spell:orb-light}
You can infuse your aether into the aetheric field of light itself, creating a condensed orb of luminosity. This manifests as floating orb of aether that can move a distance of 15 m (range 1) each turn. The orb provides full illumination within 5 m (1 combat area) of itself and \textlf{low} light within 15 m (radius 1). This lasts up to 1 hour and is a \textlf{persistent} effect.


\subsection{Overgrowth [3]}
\label{spell:overgrowth}
Plants have their own variety of aetheric field, you can infuse aether into this field to create a resonance. You may choose a region of radius 5 m radius (1 combat area) within a range of 1 (15 m) and make plant growth explode from the ground. This region is now \textlf{rough terrain}.


\subsection{Paths in the aether [2]}
You can attune your own aether to the fields within the terrain around you, allowing effortless avoidance of obstructions. This manifestation allows the aethermancer to ignore natural forms of \textlf{rough} or \textlf{dangerous terrain} effects as well as an \textlf{edge} bonus on \textlf{stealth} and \textlf{awareness} checks while in wilderness. This lasts for 1 hour.


\subsection{Prediction [3]}
Time itself resonates with aetheric frequencies. You can use this to attempt to peer into possible futures. This manifestation takes 10 minutes to complete. Afterwards, roll 3d6 and put them to one side. In addition, make an \textlf{aethermancy} check vs \textlf{difficulty} 13. If you succeed, then at any point within the next 24 hours you may replace one 3d6 roll (made by any creature or character) with the 3d6 you set aside. On a \textlf{critical failure} the GM may instead choose when to make the roll substitution. You may only have one such set of predicted dice available at once.
\subsubsection{Augment: Forecasting [3]}
You can store 2 predictions at once.


%\subsection{Pyrokinesis [3]}
%Attuning your aether to resonate with fire and heat you can adjust the aetheric field of flames to cause them to move around.  With careful control of aether you can transport the energies of an existing fire (flame and/or just heat) wherever you wish within a range 2 (25 m). 


\subsection{Resonant motion [2]}
\label{spell:min-tele}
You infuse aether into objects so that their aetheric fields resonate with motion. This allows you to exert the force of a single hand to perform simple actions on an object visible to you within range 1 (15 m).
\subsubsection{Augment: Major resonance [3]}
The aethermancer is far more attuned to the resonance of motion, allowing the aethermancer to manipulate an object up to a maximum weight of 20 kg per point of \textlf{might}. If the object is used as a weapon, it uses the aethermancer's \textlf{aethermancy} for \textlf{aim}. The bonus \textlf{Power} of such a weapon is given by the excess lifting force of the manifestation (that not required to actually move the object):  + 1 per 10 kg excess. The \textlf{power} for grabbing objects is calculated in the same manner. If multiple objects are controlled, the \textlf{Power} bonus from excess lifting force may be divided between them at the aethermancer's discretion. This manifestation requires that the aethermancer to maintain full concentration, in combat they must spend an action point to manipulate objects with this power.


\subsection{Root weaving [3] (P)}
By creating a complex aetheric resonance between yourself and nearby plants, you can call on their assistance to entrap foes. Provided there are plants or other natural growths nearby, this can be used to make roots \textlf{grapple} or \textlf{shove} a chosen creature or object within the foliage (this can be used once per turn but costs no action points). The roots use your \textlf{aethermancy} power score when making opposed checks. This is a \textlf{persistent} effect that can be maintained for up to 10 minutes.
\subsubsection{Augment: Bushwhack [3]}
Plants and roots can be made to strike at your foes, using your \textlf{aethermancy} for \textlf{aim}. For an action point you can activate a tree to strike at a nearby foe, this attack has crushing \textlf{lethality}.


\subsection{Sapping aether [4] (P)}
You infuse a discordant aetheric field into a target, weakening their body. If a target within range 1 (15 m) fails a \textlf{resist(R)} check then, whenever they lose \textlf{endurance} they lose 1 additional point. This lasts for up to 1 minute and  is a \textlf{persistent} effect.
\subsubsection{Augment: Field [2]}
If used for 2 action points this affects all creatures within a radius of 5 m (1 combat area).


\subsection{Shadows in the aether [3]}
Light has its own aetheric field and your aether can be used to create a counter vibrating field that extinguishes lights. All light disappears within a chosen radius of 1 (around 15 m). Creatures within the region fight using the \textlf{dark} lighting rules. This lasts until you cancel it.


\subsection{Transfix [3] (P)}
You infuse your aether into a creature, disrupting its aetheric field to fill its mind with a deluge of sensory information. A single target within range 1 (15 m) must \textlf{resist(R)} or stand transfixed (no actions allowed) for 10 minutes. A \textlf{resist} attempt may be made at the end of each turn to end the effect. This is \textlf{persistent} effect and any damage ends the effect immediately. 
\subsubsection{Augment: Field [2]}
If used for 2 action points this manifestation affects all creatures within a radius of 5 m (1 combat area).
\subsubsection{Augment: Amplification [3]}
If the victim of this effect is damaged, the manifestation lingers for 1 round before ending.


%\subsection{Sow doubts [3]}
%Your withering words instil doubt and uncertainty into a target that fails a \textlf{Resist(R)}, causing them an \textlf{edge} penalty on their skill checks. This lasts 1 minute, each round after the first the victim may re-attempt the \textlf{resist}. 
%\subsubsection{Augment: Sap will}
%Victims of your Sow doubts now also suffer an \textlf{edge} penalty to their \textlf{Resist} checks.
%\subsubsection{Augment: Air of doubt [11]}
%Sow doubts affects all chosen creatures adjacent to the primary target.


\subsection{Shape earth [3]}
The ground around you has an aetheric field that can be manipulated. A cunning flow of aether allows one to twist the very earth into any desired shape. The aethermancer may touch earth or stone and then manipulate up to 10 kg of earth, or 5 kg of stone, plus 10 kg earth or 5 kg stone per point of \textlf{wit}. 


\subsection{Sonic boom [3]}
By tuning your aether to resonate with the field of the air around you, a localised over-pressured pocket of air escapes outwards with a booming crash. This \textlf{Staggers} all creatures within a chosen area if they fail a \textlf{resist(R)} check. The chosen area must be within range 1 (15 m).
\subsubsection{Augment: Thunderous blast [2]}
This power now additionally inflicts damage with normal \textlf{lethality} on victims if they failed \textlf{resist} (\textlf{critical success} for the aethermancer on \textlf{resist} increments damage severity).


\subsection{Stone sense [2]}
\label{spell:stone-sense}
Your expert knowledge of aetheric vibrations allows you to extend your hearing through a continuous stone (or earthen) surface or structure. For this purpose, a wall of stone bricks is continuous but soil and dirt do not count (they do not conduct vibrations coherently enough). This power lasts for up to 1 hour.


\subsection{Stone skin [3]}
Manipulating the aetheric field of rock, you draw stone from the ground to clad you in armour. For 1 round damage rolls against you suffer a \textlf{lethality} downgrade (to a minimum of normal). 
\subsubsection{Augment: Earth-clad [2] (P)}
This makes \textlf{Stone skin} a \textlf{persistent} effect that lasts up to 10 minutes.


\subsection{Suggestion [3]}
A creature's own aetheric field can be infused with aether to make its mind more open and malleable. Make a single-sentence suggestion of an action to the target, on a failed \textlf{Resist(W)} check they follow the suggestion willingly. This suggestion will automatically fail if it would
be harmful to the victim themself or their friends/allies. Specify 1 additional suggestion per level of \textlf{critical success} for the aethermancer on \textlf{resist}. 		
\subsubsection{Augment: Manipulation [2]}
Your mastery of the aetheric fields in brains allows you to implant beliefs or feeling into a target via suggestion. In addition suggestion no longer fails if it would be harmful to the victims friends/allies.


%\subsection{Summon familiar}
%\label{spell:familiar}
%You summon a magical familiar to you, this is a very small creature (e.g. rat, cat, bird, lizard, frog, hedgehog, squirrel). The form of the creature is changeable and it takes 10 minutes of concentration to change its form. You can communicate telepathically with the creature while it is within 30 m of you. Additionally, you can view the sensory perceptions of the creature instead of your own while it is within 30 m of you (you are blind and deaf to your own senses while doing this). This spell occupies a passive \textlf{perk} slot.


\subsection{Time warp [5]}
Projecting your aether towards another creature, you make its own aetheric field discordant with the flow of time. Choose a creature within range 1 (15 m) to either gain or lose an action point each turn until the effect ends (maximum duration 1 minute). Targets may \textlf{resist(R)} each turn to end the effect. Only one action point may be gained or lost each turn in this way. \textlf{critical success} for the aethermancer on \textlf{resist} means the target cannot attempt \textlf{resist} next turn.
\subsubsection{Augment: Field [2]}
If used for 2 action points this affects all creatures within a targeted radius of 5 m (1 combat area).


\subsection{Vortex [3]}
Infusing a sudden burst of aether into the air creates a resonance that roars out as a powerful vortex of wind. This reverses the flight of projectiles entering a target region within range 1 (15 m) for 1 round. 
\subsubsection{Augment: Howling winds [2]}
Your mastery of the air means that this power can be  manifested with ferocious winds that knock down all creatures in the target area if they fail a \textlf{resist(M)} check.


\subsection{Warp space [5]}
You infuse your aether around a creature, warping the aetheric field of space itself. For 1 action point you can move a target creature, which can attempt to \textlf{resist(C)} if it wants to. This moves the target a distance up to range 2 (25 m). This cannot affect targets more than 1 size category larger than the aethermancer without 1 level of \textlf{critical success} for the aethermancer on \textlf{resist} per size additional category difference.


\subsection{Weather sense [2]}
You attune to the aetheric fields of air and water, thus you can predict the weather up to 1 day + \textlf{wit} in advance.
\subsubsection{Augment: Weather touch [4]}
Your mastery of air and water allows you to nudge the weather. This lets you alter the weather, with the effect occurring after 5 - X days where X is the number of aether points spent on this manifestation. 